---
title: "Neural Networks"
subtitle: "Homework 2 - Data Science 2 Course"
author: "Kata Süle"
date: '5th April 2021'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

```{r, include=FALSE}
# import packages
library(keras)
library(ggplot2)
library(kableExtra)
library(knitr)
library(grid)
library(magick)
library(filesstrings)

# set path
path <- 'C:/CEU/Winter_Term/Data_Science_2/Homework 2/'
```


# 1. Fashion MNIST data

### a. Show some example images from the data.

The train dataset contains 60,000, while the test contains 10,000 grayscale images of fashion items. The items belong to 10 classes which are indicated by numbers ranging from 0 to 9. The data is stored in three dimensional arrays. Below are the first 6 fashion items extracted from the train dataset. The numbers below the pictures indicate the classes they belong to.

```{r, include=FALSE, cache=TRUE}
# import data
fashion_mnist <- dataset_fashion_mnist()
x_train <- fashion_mnist$train$x
y_train <- fashion_mnist$train$y
x_test <- fashion_mnist$test$x
y_test <- fashion_mnist$test$y
```

```{r}
# split the canvas into rows and columns
par(mfrow = c(2, 3))

# plot the first 6 images
for (i in 1:6){
  image(
    x_train[i,,],
    col = gray.colors(255), xlab = y_train[i], ylab = ""
  )
}
```

### b. Train a fully connected deep network to predict items.

In order to be able to train a fully connected neural network the data had to be restructured. First, I split the original training set into training and validation sets with 10,000 and 50,000 observations respectively. After this step I rescaled the data by dividing it by 255 because for grayscale images the value of pixels is between 0 and 255. Furthermore, I also converted it to two dimensional matrices so that the models can be trained on it later. As for the labels I created one-hot encoding for them.

```{r, include=FALSE, cache=TRUE}
### RESTRUCTURE DATA
# split train data into train and validation sets
# get indeces
set.seed(1234)
train_indices <- sample(seq(nrow(x_train)), 10000) # keep 10000 observations in the training set

# split x into train and validation
data_train_x <- x_train[train_indices,, ]
data_valid_x <- x_train[-train_indices,, ]

# split y into train and validation
data_train_y <- y_train[train_indices]
data_valid_y <- y_train[-train_indices]

# rescale x for all sets
data_train_x <- as.matrix(as.data.frame.array(data_train_x)) / 255
data_valid_x <- as.matrix(as.data.frame.array(data_valid_x)) / 255
data_test_x <- as.matrix(as.data.frame.array(x_test)) / 255

# one-hot encoding for y for all sets
data_train_y <- to_categorical(data_train_y, 10)
data_valid_y <- to_categorical(data_valid_y, 10)
data_test_y <- to_categorical(y_test, 10)
```

In total I trained a baseline model based on the script that we used in class and four additional models with different settings. There were always 784 input nodes because of the size of the images (28x28 pixels) and 10 output nodes because there were 10 classes (0-9). The further settings of the models were the following:

1. Baseline model:

  * 1 hidden layer with 128 nodes
  * dropout rate of 0.3
  * relu and softmax activation functions.
  
2. Model 1:

  * 1 hidden layer with 28 nodes
  * relu and softmax activation functions.
  
3. Model 2:

  * 2 hidden layers with 28 and 28 nodes
  * relu, sigmoid and softmax activation functions.
  
4. Model 3:

  * 3 hidden layers with 100 nodes in each
  * dropout rate of 0.3
  * relu, sigmoid, tanh and softmax activation functions.
  
5. Model 4:

  * 2 hidden layers with 100 nodes each
  * dropout rate of 0.7
  * sigmoid, tanh and softmax activation functions.

```{r, include=FALSE}
##### TRAIN MODELS
### BASELINE MODEL
# create model object
model_base <- keras_model_sequential()

# add layers to the model
model_base %>%
  layer_dense(units = 128, activation = 'relu', input_shape = c(784)) %>%
  layer_dropout(rate = 0.3) %>%
  layer_dense(units = 10, activation = 'softmax')

# check model summary
summary(model_base)

# set loss function and the metrics we want to see
compile(
  model_base,
  loss = 'categorical_crossentropy',
  optimizer = optimizer_rmsprop(),
  metrics = c('accuracy')
)

# fit model
# baseRDS_NN <- fit(
#   model_base, data_train_x, data_train_y,
#   epochs = 30, batch_size = 128,
#   validation_data = list(data_valid_x, data_valid_y)
# )

# save fit
# saveRDS(baseRDS_NN, paste0(path, "data/baseRDS_NN.rds"))
baseRDS_NN <- readRDS(paste0(path, "data/baseRDS_NN.rds") )

# save model
# save_model_hdf5(model_base, paste0(path, "data/model_base.h5"))

# import model
model_base <- load_model_hdf5(paste0(path, "data/model_base.h5"))

# evaluate model
model_base_eval <- as.data.frame(evaluate(model_base, data_valid_x, data_valid_y))
```

```{r, include=FALSE}
### FIRST MODEL
# create model object
model1 <- keras_model_sequential()

# add layers to the model
model1 %>%
  layer_dense(units = 28, activation = 'relu', input_shape = c(784)) %>% 
  layer_dense(units = 10, activation = 'softmax')

# check model summary
summary(model1)

# set loss function and the metrics we want to see
compile(
  model1,
  loss = 'categorical_crossentropy',
  optimizer = optimizer_rmsprop(),
  metrics = c('accuracy')
)

# fit model
# model1RDS_NN <- fit(
#   model1, data_train_x, data_train_y,
#   epochs = 30, batch_size = 128,
#   validation_data = list(data_valid_x, data_valid_y)
# )

# save fit
# saveRDS(model1RDS_NN, paste0(path, "data/model1RDS_NN.rds") )
model1RDS_NN <- readRDS(paste0(path, "data/model1RDS_NN.rds"))

# save model
# save_model_hdf5(model1, paste0(path, "data/model1.h5"))

# import model
model1 <- load_model_hdf5(paste0(path, "data/model1.h5"))

# evaluate model
model1_eval <- as.data.frame(evaluate(model1, data_valid_x, data_valid_y))
```

```{r, include=FALSE}
### SECOND MODEL
# create model object
model2 <- keras_model_sequential()

# add layers to the model
model2 %>%
  layer_dense(units = 28, activation = 'relu', input_shape = c(784)) %>%
  layer_dense(units = 28, activation = 'sigmoid') %>%
  layer_dense(units = 10, activation = 'softmax')

# check model summary
summary(model2)

# set loss function and the metrics we want to see
compile(
  model2,
  loss = 'categorical_crossentropy',
  optimizer = optimizer_rmsprop(),
  metrics = c('accuracy')
)

# fit model
# model2RDS_NN <- fit(
#   model2, data_train_x, data_train_y,
#   epochs = 30, batch_size = 128,
#   validation_data = list(data_valid_x, data_valid_y)
# )

# save fit
# saveRDS(model2RDS_NN, paste0(path, "data/model2RDS_NN.rds") )
model2RDS_NN <- readRDS(paste0(path, "data/model2RDS_NN.rds"))

# save model
# save_model_hdf5(model2, paste0(path, "data/model2.h5"))

# import model
model2 <- load_model_hdf5(paste0(path, "data/model2.h5"))

# evaluate model
model2_eval <- as.data.frame(evaluate(model2, data_valid_x, data_valid_y))
```

```{r, include=FALSE}
### THIRD MODEL
# create model object
model3 <- keras_model_sequential()

# add layers to the model
model3 %>%
  layer_dense(units = 100, activation = 'relu', input_shape = c(784)) %>%
  layer_dense(units = 100, activation = 'sigmoid') %>%
  layer_dense(units = 100, activation = 'tanh') %>%
  layer_dropout(rate = 0.3) %>% 
  layer_dense(units = 10, activation = 'softmax')

# check model summary
summary(model3)

# set loss function and the metrics we want to see
compile(
  model3,
  loss = 'categorical_crossentropy',
  optimizer = optimizer_rmsprop(),
  metrics = c('accuracy')
)

# fit model
# model3RDS_NN <- fit(
#   model3, data_train_x, data_train_y,
#   epochs = 30, batch_size = 128,
#   validation_data = list(data_valid_x, data_valid_y)
# )

# save fit
# saveRDS(model3RDS_NN, paste0(path, "data/model3RDS_NN.rds"))
model3RDS_NN <- readRDS(paste0(path, "data/model3RDS_NN.rds"))

# save model
# save_model_hdf5(model3, paste0(path, "data/model3.h5"))

# import model
model3 <- load_model_hdf5(paste0(path, "data/model3.h5"))

# evaluate model
model3_eval <- as.data.frame(evaluate(model3, data_valid_x, data_valid_y))
```

```{r, include=FALSE}
### FOURTH MODEL
# create model object
model4 <- keras_model_sequential()

# add layers to the model
model4 %>%
  layer_dense(units = 100, activation = 'sigmoid', input_shape = c(784)) %>%
  layer_dense(units = 100, activation = 'tanh') %>%
  layer_dropout(rate = 0.7) %>% 
  layer_dense(units = 10, activation = 'softmax')

# check model summary
summary(model4)

# set loss function and the metrics we want to see
compile(
  model4,
  loss = 'categorical_crossentropy',
  optimizer = optimizer_rmsprop(),
  metrics = c('accuracy')
)

# fit model
# model4RDS_NN <- fit(
#   model4, data_train_x, data_train_y,
#   epochs = 30, batch_size = 128,
#   validation_data = list(data_valid_x, data_valid_y)
# )

# save fit
# saveRDS(model4RDS_NN, paste0(path, "data/model4RDS_NN.rds"))
model4RDS_NN <- readRDS(paste0(path, "data/model4RDS_NN.rds"))

# save model
# save_model_hdf5(model4, paste0(path, "data/model4.h5"))

# import model
model4 <- load_model_hdf5(paste0(path, "data/model4.h5"))

# evaluate model
model4_eval <- as.data.frame(evaluate(model4, data_valid_x, data_valid_y))
```

The table below shows the accuracy and loss measures for all the models calculated for the validation set. We can see that the baseline model proved to have the best performance. Increasing the number of hidden layers usually resulted in better performance so did setting a dropout rate. Furthermore, including more nodes in the hidden layers seemed to matter as well. As for the activation functions I did not observe a large difference between the different settings.

```{r}
# model comparison
# bind the data frames
model_comparison <- cbind(model_base_eval, model1_eval, model2_eval, model3_eval, model4_eval)

# rename columns
colnames(model_comparison) <- c('Baseline', 'Model 1', 'Model 2', 'Model 3', 'Model 4')

# create table
knitr::kable(model_comparison, caption = 'Model comparison of fully connected neural networks', digits = 2) %>% 
  kable_styling( position = "center", latex_options = 'hold_position', bootstrap_options = c("striped", "hover"))
```

The plot below shows the training history of the baseline model which I chose to be the final model. We can see how the performance stabilizes over the epochs.

```{r}
# plot final model
plot(baseRDS_NN) +
  ggtitle('Training history of final model') +
  theme_bw() +
  xlab('Epoch') +
  ylab('')
```

### c. Evaluate the model on the test set. How does test error compare to validation error?

The table below shows the performance of the final model (baseline model) on the training, validation and test sets. We can see that the test error is not very different from the validation error, therefore the model has a relatively stable performance.

```{r}
# calculate test set performance of baseline model
model_base_eval_test <- as.data.frame(evaluate(model_base, data_test_x, data_test_y))
model_base_eval_train <- as.data.frame(evaluate(model_base, data_train_x, data_train_y))

# bind performances on different sets
final_model_comparison <- cbind(model_base_eval_train, model_base_eval, model_base_eval_test)

# rename rows
colnames(final_model_comparison) <- c('Training set', 'Validation set', 'Test set')

# create table
knitr::kable(final_model_comparison, caption = 'Performance of final model on different sets', digits = 2) %>% 
  kable_styling( position = "center", latex_options = 'hold_position', bootstrap_options = c("striped", "hover"))
```

### d. Try building a convolutional neural network and see if you can improve test set performance.

In this part I trained a baseline model based on the script that we used in class and five additional models. The input shape (28, 28, 1) and the number of output nodes (10) were the same across the models. The differences between the models were the following: kernel size, pool size, activation functions, number of nodes in the hidden layers, dropout rate and number of convolutional layers. Based on the table below we can say that again the baseline model had the best performance on the validation set. In general larger kernel size (5x5 instead of 3x3) decreased performance so did adding too many convolutional layers. On the other hand adding more hidden layers after the convolutional ones helped to increase performance.

```{r, include=FALSE}
# reshape data to be able to train models
data_train_x <- array_reshape(data_train_x, c(nrow(data_train_x), 28, 28, 1))
data_valid_x <- array_reshape(data_valid_x, c(nrow(data_valid_x), 28, 28, 1))
data_test_x <- array_reshape(data_test_x, c(nrow(data_test_x), 28, 28, 1))
```

```{r, include=FALSE}
### TRAIN MODELS
### BASELINE MODEL

# create model object
cnn_model_base <- keras_model_sequential()

# add layers
cnn_model_base %>%
  layer_conv_2d(
    filters = 32,
    kernel_size = c(3, 3),
    activation = 'relu',
    input_shape = c(28, 28, 1)
  ) %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_dropout(rate = 0.25) %>%
  layer_flatten() %>% # increases complexity
  layer_dense(units = 16, activation = 'relu') %>% 
  layer_dense(units = 10, activation = 'softmax') 

# compile model
compile(
  cnn_model_base,
  loss = 'categorical_crossentropy',
  optimizer = optimizer_rmsprop(),
  metrics = c('accuracy')
)

# fit model
# baseRDS_CNN <- fit(
#   cnn_model_base, data_train_x, data_train_y,
#   epochs = 15, batch_size = 128,
#   validation_data = list(data_valid_x, data_valid_y)
# )

# save fit
# saveRDS(baseRDS_CNN, paste0(path, "data/baseRDS_CNN.rds"))
baseRDS_CNN <- readRDS(paste0(path, "data/baseRDS_CNN.rds"))

# save model
# save_model_hdf5(cnn_model_base, paste0(path, "data/cnn_model_base.h5"))

# import model
cnn_model_base <- load_model_hdf5(paste0(path, "data/cnn_model_base.h5"))

# evaluate model
model_base_cnn_eval <- as.data.frame(evaluate(cnn_model_base, data_valid_x, data_valid_y))
```

```{r, include=FALSE}
### CNN Model 1

# create model object
cnn_model1 <- keras_model_sequential()

# add layers
cnn_model1 %>%
  layer_conv_2d(
    filters = 10,
    kernel_size = c(5, 5), 
    activation = 'relu',
    input_shape = c(28, 28, 1)
  ) %>%
  layer_max_pooling_2d(pool_size = c(3, 3)) %>% 
  layer_dropout(rate = 0.25) %>%
  layer_flatten() %>% 
  layer_dense(units = 10, activation = 'relu') %>% 
  layer_dense(units = 10, activation = 'softmax') 

# compile model
compile(
  cnn_model1,
  loss = 'categorical_crossentropy',
  optimizer = optimizer_rmsprop(),
  metrics = c('accuracy')
)

# fit model
# model1RDS_CNN <- fit(
#   cnn_model1, data_train_x, data_train_y,
#   epochs = 15, batch_size = 128,
#   validation_data = list(data_valid_x, data_valid_y)
# )

# save fit
# saveRDS(model1RDS_CNN, paste0(path, "data/model1RDS_CNN.rds"))
model1RDS_CNN <- readRDS(paste0(path, "data/model1RDS_CNN.rds"))

# save model
# save_model_hdf5(cnn_model1, paste0(path, "data/cnn_model1.h5"))

# import model
cnn_model1 <- load_model_hdf5(paste0(path, "data/cnn_model1.h5"))

# evaluate model
model1_cnn_eval <- as.data.frame(evaluate(cnn_model1, data_valid_x, data_valid_y))
```

```{r, include=FALSE}
### CNN Model 2

# create model object
cnn_model2 <- keras_model_sequential()

# add layers
cnn_model2 %>%
  layer_conv_2d(
    filters = 10, 
    kernel_size = c(5, 5), 
    activation = 'relu',
    input_shape = c(28, 28, 1)
  ) %>%
  layer_max_pooling_2d(pool_size = c(3, 3)) %>% 
  layer_conv_2d(
    filters = 20, 
    kernel_size = c(5, 5),
    activation = 'relu',
    input_shape = c(28, 28, 1)
  ) %>%
  layer_max_pooling_2d(pool_size = c(3, 3)) %>% 
  layer_dropout(rate = 0.25) %>%
  layer_flatten() %>% 
  layer_dense(units = 10, activation = 'relu') %>% 
  layer_dense(units = 10, activation = 'softmax') 

# compile model
compile(
  cnn_model2,
  loss = 'categorical_crossentropy',
  optimizer = optimizer_rmsprop(),
  metrics = c('accuracy')
)

# fit model
# model2RDS_CNN <- fit(
#   cnn_model2, data_train_x, data_train_y,
#   epochs = 15, batch_size = 128,
#   validation_data = list(data_valid_x, data_valid_y)
# )

# save fit
# saveRDS(model2RDS_CNN, paste0(path, "data/model2RDS_CNN.rds"))
model2RDS_CNN <- readRDS(paste0(path, "data/model2RDS_CNN.rds"))

# save model
# save_model_hdf5(cnn_model2, paste0(path, "data/cnn_model2.h5"))

# import model
cnn_model2 <- load_model_hdf5(paste0(path, "data/cnn_model2.h5"))

# evaluate model
model2_cnn_eval <- as.data.frame(evaluate(cnn_model2, data_valid_x, data_valid_y))
```

```{r, include=FALSE}
### CNN Model 3

# create model object
cnn_model3 <- keras_model_sequential()

# add layers
cnn_model3 %>% 
  layer_conv_2d(filters = 32, kernel_size = c(3,3), activation = "relu", 
                input_shape = c(28,28,1)) %>% 
  layer_max_pooling_2d(pool_size = c(2,2)) %>% 
  layer_conv_2d(filters = 64, kernel_size = c(3,3), activation = "relu") %>% 
  layer_max_pooling_2d(pool_size = c(2,2)) %>% 
  layer_conv_2d(filters = 64, kernel_size = c(3,3), activation = "relu") %>% 
  layer_dropout(rate = 0.25) %>%
  layer_flatten() %>% 
  layer_dense(units = 10, activation = 'relu') %>%
  layer_dense(units = 10, activation = 'softmax') 

# compile model
compile(
  cnn_model3,
  loss = 'categorical_crossentropy',
  optimizer = optimizer_rmsprop(),
  metrics = c('accuracy')
)

# fit model
# model3RDS_CNN <- fit(
#   cnn_model3, data_train_x, data_train_y,
#   epochs = 15, batch_size = 128,
#   validation_data = list(data_valid_x, data_valid_y)
# )

# save fit
# saveRDS(model3RDS_CNN, paste0(path, "data/model3RDS_CNN.rds"))
model3RDS_CNN <- readRDS(paste0(path, "data/model3RDS_CNN.rds"))

# save model
# save_model_hdf5(cnn_model3, paste0(path, "data/cnn_model3.h5"))

# import model
cnn_model3 <- load_model_hdf5(paste0(path, "data/cnn_model3.h5"))

# evaluate model
model3_cnn_eval <- as.data.frame(evaluate(cnn_model3, data_valid_x, data_valid_y))
```

```{r, include=FALSE}
### CNN Model 4

# create model object
cnn_model4 <- keras_model_sequential()

# add layers
cnn_model4 %>% 
  layer_conv_2d(filters = 32, kernel_size = c(3,3), activation = "sigmoid", 
                input_shape = c(28,28,1)) %>% 
  layer_max_pooling_2d(pool_size = c(2,2)) %>% 
  layer_dropout(rate = 0.25) %>%
  layer_flatten() %>% 
  layer_dense(units = 10, activation = 'softmax')

# compile model
compile(
  cnn_model4,
  loss = 'categorical_crossentropy',
  optimizer = optimizer_rmsprop(),
  metrics = c('accuracy')
)

# fit model
# model4RDS_CNN <- fit(
#   cnn_model4, data_train_x, data_train_y,
#   epochs = 15, batch_size = 128,
#   validation_data = list(data_valid_x, data_valid_y)
# )

# save fit
# saveRDS(model4RDS_CNN, paste0(path, "data/model4RDS_CNN.rds"))
model4RDS_CNN <- readRDS(paste0(path, "data/model4RDS_CNN.rds"))

# save model
# save_model_hdf5(cnn_model4, paste0(path, "data/cnn_model4.h5"))

# import model
cnn_model4 <- load_model_hdf5(paste0(path, "data/cnn_model4.h5"))

# evaluate model
model4_cnn_eval <- as.data.frame(evaluate(cnn_model4, data_valid_x, data_valid_y))

```

```{r, include=FALSE}
### CNN Model 5

# create model object
cnn_model5 <- keras_model_sequential()

# add layers
cnn_model5 %>% 
  layer_conv_2d(filters = 32, kernel_size = c(3,3), activation = "relu", 
                input_shape = c(28,28,1)) %>% 
  layer_max_pooling_2d(pool_size = c(2,2)) %>% 
  layer_dropout(rate = 0.25) %>%
  layer_flatten() %>% # increases complexity
  layer_dense(units = 32, activation = 'relu') %>%  
  layer_dense(units = 32, activation = 'sigmoid') %>%  
  layer_dense(units = 10, activation = 'softmax') 

# compile model
compile(
  cnn_model5,
  loss = 'categorical_crossentropy',
  optimizer = optimizer_rmsprop(),
  metrics = c('accuracy')
)

# fit model
# model5RDS_CNN <- fit(
#   cnn_model5, data_train_x, data_train_y,
#   epochs = 15, batch_size = 128,
#   validation_data = list(data_valid_x, data_valid_y)
# )

# save fit
# saveRDS(model5RDS_CNN, paste0(path, "data/model5RDS_CNN.rds"))
model5RDS_CNN <- readRDS(paste0(path, "data/model5RDS_CNN.rds"))

# save model
# save_model_hdf5(cnn_model5, paste0(path, "data/cnn_model5.h5"))

# import model
cnn_model5 <- load_model_hdf5(paste0(path, "data/cnn_model5.h5"))

# evaluate model
model5_cnn_eval <- as.data.frame(evaluate(cnn_model5, data_valid_x, data_valid_y))
```

```{r}
# model comparison
# bind model performances
model_comparison_cnn <- cbind(model_base_cnn_eval, model1_cnn_eval, model2_cnn_eval, model3_cnn_eval, model4_cnn_eval, model5_cnn_eval)

# rename rows
colnames(model_comparison_cnn) <- c('Baseline', 'Model 1', 'Model 2', 'Model 3', 'Model 4', 'Model 5')

# create table
knitr::kable(model_comparison_cnn, caption = 'Model comparison for convolutional networks', digits = 2) %>% 
  kable_styling( position = "center", latex_options = 'hold_position', bootstrap_options = c("striped", "hover"))
```

The plot below shows the training history of the best model (baseline model). We can see the signs of overfitting between the 12 and 15 epoch.

```{r}
# plot best model
plot(baseRDS_CNN) +
  ggtitle('Training history of final convolutional model') +
  theme_bw() +
  xlab('Epoch') +
  ylab('')
```

# 2. Hot dog or not hot dog?

### a. Pre-process data so that it is acceptable by Keras (set folder structure, bring images to the same size, etc).

First I created a validation set by randomly selecting 150-150 photos from the hot_dog and not_hot_dog folders of the training data. Then I rescaled the images, set the folders for the train, validation and test generator functions and set the image and batch sizes as well.

```{r, include=FALSE}
# create validation set

# randomly sample 150 photos from the hot dog and from the not hot dog training set folders
# set.seed(1234)
# validation_indeces_yes <- sample(list.files(paste0(path, "data/hot-dog-not-hot-dog/train/hot_dog")), size = 150, replace = F)
# validation_indeces_no <- sample(list.files(paste0(path, "data/hot-dog-not-hot-dog/train/not_hot_dog")), size = 150, replace = F)

# create new folders for the validation set
# dir.create(paste0(path,"data/hot-dog-not-hot-dog/validation"))
# dir.create(paste0(path,"data/hot-dog-not-hot-dog/validation/hot_dog"))
# dir.create(paste0(path,"data/hot-dog-not-hot-dog/validation/not_hot_dog"))

# move hot dog photos from train to validation if they are in the sample
# for (file in list.files(paste0(path, "data/hot-dog-not-hot-dog/train/hot_dog"))) {
#   if (file %in% validation_indeces_yes){
#     file.move(paste0(path, "data/hot-dog-not-hot-dog/train/hot_dog/", file), paste0(path,"data/hot-dog-not-hot-dog/validation/hot_dog/"))
#   }
# }

# move not hot dog photos from train to validation if they are in the sample
# for (file in list.files(paste0(path, "data/hot-dog-not-hot-dog/train/not_hot_dog"))) {
#   if (file %in% validation_indeces_no){
#     file.move(paste0(path, "data/hot-dog-not-hot-dog/train/not_hot_dog/", file), paste0(path,"data/hot-dog-not-hot-dog/validation/not_hot_dog/"))
#   }
# }
```

```{r, include=FALSE}
# rescale images
train_datagen <- image_data_generator(rescale = 1/255)  
validation_datagen <- image_data_generator(rescale = 1/255)  
test_datagen <- image_data_generator(rescale = 1/255) 
```

```{r, include=FALSE}
# set image and batch size
image_size <- c(150, 150)
batch_size <- 50

# set train generator
train_generator <- flow_images_from_directory(
  file.path(path, "data/hot-dog-not-hot-dog/train"), 
  train_datagen,              
  target_size = image_size,  
  batch_size = batch_size,
  class_mode = "binary"      
)

# set validation generator
validation_generator <- flow_images_from_directory(
  file.path(path, "data/hot-dog-not-hot-dog/validation"),   
  validation_datagen,
  target_size = image_size,
  batch_size = batch_size,
  class_mode = "binary"
)

# set test generator
test_generator <- flow_images_from_directory(
  file.path(path, "data/hot-dog-not-hot-dog/test"), 
  test_datagen,
  target_size = image_size,
  batch_size = batch_size,
  class_mode = "binary"
)
```

### b. Estimate a convolutional neural network to predict if an image contains a hot dog or not. Evaluate your model on the test set.

I estimated a baseline model which had two convolutional layers, one hidden layer and a dropout rate of 0.5. The plot below shows the training history of this model. The plot shows sign of overfitting because the training accuracy keeps increasing when the validation accuracy starts to decrease.

```{r, include=FALSE}
### TRAIN MODELS
### BASELINE MODEL

# create model object and add layers
hot_dog_model_baseline <- keras_model_sequential()%>%
  layer_conv_2d(filters = 32,
                kernel_size = c(3, 3), 
                activation = 'relu',
                input_shape = c(150, 150, 3)) %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_conv_2d(filters = 64,
                kernel_size = c(3, 3),
                activation = 'relu') %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_dropout(rate = 0.5) %>% 
  layer_flatten() %>% 
  layer_dense(units = 64, activation = 'relu') %>% 
  layer_dropout(rate = 0.5) %>%
  layer_dense(units = 1, activation = "sigmoid")  

# compile model
hot_dog_model_baseline %>% compile(
  loss = "binary_crossentropy",
  optimizer = 'rmsprop',
  metrics = c("accuracy")
)

# fit model
# hot_dog_baseline_RDS <- hot_dog_model_baseline %>% fit(
#   train_generator,
#   steps_per_epoch = 198 / batch_size, # divide size of training dataset by batch size
#   epochs = 20,
#   validation_data = validation_generator,
#   validation_steps = 300 / batch_size # divide size of training dataset by batch size
# )

# save fit 
# saveRDS(hot_dog_baseline_RDS, paste0(path, "data/hot_dog_baseline_RDS.rds"))
hot_dog_baseline_RDS <- readRDS(paste0(path, "data/hot_dog_baseline_RDS.rds"))

# save model
# save_model_hdf5(hot_dog_model_baseline, paste0(path, "data/hot_dog_model_baseline.h5"))

# Recreate the exact same model purely from the file
hot_dog_model_baseline <- load_model_hdf5(paste0(path, "data/hot_dog_model_baseline.h5"))

# evaluate model
hot_dog_model_baseline_eval <- as.data.frame(evaluate_generator(hot_dog_model_baseline, test_generator, steps = 10))
```

```{r}
# plot training history of model
plot(hot_dog_baseline_RDS) +
  ggtitle('Training history of baseline convolutional image classifier') +
  theme_bw() +
  xlab('Epoch') +
  ylab('')
```

The table below shows the test set performance of the baseline model which is close to a coin flip.

```{r}
# rename columns
colnames(hot_dog_model_baseline_eval) <- 'Baseline model'

# show test set performance
knitr::kable(hot_dog_model_baseline_eval, caption = 'Test set performance of baseline convolutional image classifier') %>% 
  kable_styling( position = "center", latex_options = 'hold_position', bootstrap_options = c("striped", "hover"))
```

Besides the baseline model I trained another model with four convolutional layers. The plot below shows its training history. Unlike the one before this does not show signs of overfitting.

```{r, include=FALSE}
### FINAL MODEL

# create model object and add layers
hot_dog_model_final <- keras_model_sequential()%>%
  layer_conv_2d(filters = 32,
                kernel_size = c(3, 3), 
                activation = 'relu',
                input_shape = c(150, 150, 3)) %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_conv_2d(filters = 32,
                kernel_size = c(3, 3), 
                activation = 'relu') %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_conv_2d(filters = 64,
                kernel_size = c(3, 3), 
                activation = 'relu') %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_conv_2d(filters = 16,
                kernel_size = c(3, 3), 
                activation = 'relu') %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_dropout(rate = 0.5) %>% 
  layer_flatten() %>% 
  layer_dense(units = 64, activation = 'relu') %>% 
  layer_dropout(rate = 0.5) %>%
  layer_dense(units = 1, activation = "sigmoid")   # for binary

# compile model
hot_dog_model_final %>% compile(
  loss = "binary_crossentropy",
  optimizer = 'rmsprop',
  metrics = c("accuracy")
)

# fit model
# hot_dog_final_RDS <- hot_dog_model_final %>% fit(
#   train_generator,
#   steps_per_epoch = 198 / batch_size, # divide size of training dataset by batch size
#   epochs = 20,
#   validation_data = validation_generator,
#   validation_steps = 300 / batch_size, # divide size of training dataset by batch size
#   callback_reduce_lr_on_plateau(monitor = "val_loss", factor = 0.1) # if crossentropy loss plateaus, change LR
# )

# save fit 
# saveRDS(hot_dog_final_RDS, paste0(path, "data/hot_dog_final_RDS.rds"))
hot_dog_final_RDS <- readRDS(paste0(path, "data/hot_dog_final_RDS.rds"))

# save model
# save_model_hdf5(hot_dog_model_final, paste0(path, "data/hot_dog_model_final.h5"))

# import model
hot_dog_model_final <- load_model_hdf5(paste0(path, "data/hot_dog_model_final.h5"))

# evaluate model
hot_dog_final_model_eval <- as.data.frame(evaluate_generator(hot_dog_model_final, test_generator, steps = 10))
```

```{r}
# plot training history
plot(hot_dog_final_RDS) +
  ggtitle('Training history of final convolutional image classifier') +
  theme_bw() +
  xlab('Epoch') +
  ylab('')
```

The table below shows the test set performance of this second model. We can conclude that the performances of both models are very poor.

```{r}
# rename columns
colnames(hot_dog_final_model_eval) <- 'Final model'

# show test set performance
knitr::kable(hot_dog_final_model_eval, caption = 'Test set performance of final convolutional image classifier') %>% 
  kable_styling( position = "center", latex_options = 'hold_position', bootstrap_options = c("striped", "hover"))
```

### c. Could data augmentation techniques help with achieving higher predictive accuracy?

Since the accuracy of the first two models was very poor I experimented with augmentation techniques to improve it. I estimated three models in total with different augmentation techniques. I applied rotation with varying angles, width and height shift, zooming, changing the brightness, flipping the images and setting the fill_mode to nearest so that empty pixels would be coloured based on the nearest pixels.

The table below shows the validation set performances of the augmented models versus the final model chosen in the previous exercise. A larger angle for rotation and a larger value for zoom increased performance the most. The zoom parameter probably helped a lot because the hot dogs in many of the photos only account for a relatively small part of the picture.

```{r, include=FALSE}
# list to save model performances into
model_eval_list <- list()

# add validation performance of not augmented final model
model_eval_list['final_model_no_augmentation'] <- as.data.frame(evaluate_generator(hot_dog_model_final, validation_generator, steps = 10))
```

```{r, include=FALSE}
### FIRST DATA AUGMENTATION
# set augmentation parameters
train_datagen = image_data_generator(
  rescale = 1/255,
  rotation_range = 40,
  width_shift_range = 0.2,
  height_shift_range = 0.2,
  shear_range = 0.2,
  zoom_range = 0.2, 
  horizontal_flip = TRUE,
  fill_mode = "nearest"
)

# reinitialize training generator
train_generator <- flow_images_from_directory(
  file.path(path, "data/hot-dog-not-hot-dog/train"),  
  train_datagen,              
  target_size = image_size,  
  batch_size = batch_size,
  class_mode = "binary"       
)

# fit model
# hot_dog_final_aug1_RDS <- hot_dog_model_final %>% fit(
#   train_generator,
#   steps_per_epoch = 198 / batch_size, # divide size of training dataset by batch size
#   epochs = 20,
#   validation_data = validation_generator,
#   validation_steps = 300 / batch_size, # divide size of training dataset by batch size
#   callback_reduce_lr_on_plateau(monitor = "val_loss", factor = 0.1) # if crossentropy loss plateaus, change LR
# )

# save fit
# saveRDS(hot_dog_final_aug1_RDS, paste0(path,"data/hot_dog_final_aug1_RDS.rds"))
hot_dog_final_aug1_RDS <- readRDS(paste0(path,"data/hot_dog_final_aug1_RDS.rds"))

# save model
# save_model_hdf5(hot_dog_model_final, paste0(path, "data/hot_dog_model_aug1.h5"))

# import model
hot_dog_model_final <- load_model_hdf5(paste0(path, "data/hot_dog_model_aug1.h5"))

# add performance to list
model_eval_list['first_try_augmentation'] <- as.data.frame(evaluate_generator(hot_dog_model_final, validation_generator, steps = 10))
```

```{r, include=FALSE}
### SECOND DATA AUGMENTATION

# set augmentation parameters
train_datagen = image_data_generator(
  rescale = 1/255,
  rotation_range = 90,
  width_shift_range = 0.4,
  height_shift_range = 0.4,
  zoom_range = 0.5, 
  brightness_range = c(1,1.5),
  fill_mode = "nearest" 
)

# reinitialize training generator
train_generator <- flow_images_from_directory(
  file.path(path, "data/hot-dog-not-hot-dog/train"), 
  train_datagen,              
  target_size = image_size,  
  batch_size = batch_size,
  class_mode = "binary"       
)

# fit model
# hot_dog_final_aug2_RDS <- hot_dog_model_final %>% fit(
#   train_generator,
#   steps_per_epoch = 198 / batch_size, # divide size of training dataset by batch size
#   epochs = 20,
#   validation_data = validation_generator,
#   validation_steps = 300 / batch_size, # divide size of training dataset by batch size
#   callback_reduce_lr_on_plateau(monitor = "val_loss", factor = 0.1) # if crossentropy loss plateaus, change LR
# )

# save fit
# saveRDS(hot_dog_final_aug2_RDS, paste0(path, "data/hot_dog_final_aug2_RDS.rds"))
hot_dog_final_aug2_RDS <- readRDS(paste0(path, "data/hot_dog_final_aug2_RDS.rds"))

# save model
# save_model_hdf5(hot_dog_model_final, paste0(path, "data/hot_dog_model_aug2.h5"))

# import model
hot_dog_model_final <- load_model_hdf5(paste0(path, "data/hot_dog_model_aug2.h5"))

# add performance to list
model_eval_list['second_try_augmentation'] <- as.data.frame(evaluate_generator(hot_dog_model_final, validation_generator, steps = 10))
```

```{r, include=FALSE}
### THIRD DATA AUGMENTATION

# set augmentation parameters
train_datagen = image_data_generator(
  rescale = 1/255,
  rotation_range = 10,
  width_shift_range = 0.5,
  height_shift_range = 0.5,
  zoom_range = 0.1, 
  brightness_range = c(1.5,2),
  fill_mode = "wrap" 
)

# reinitialize training generator
train_generator <- flow_images_from_directory(
  file.path(path, "data/hot-dog-not-hot-dog/train"), 
  train_datagen,              
  target_size = image_size,  
  batch_size = batch_size,
  class_mode = "binary"       
)

# fit model
# hot_dog_final_aug3_RDS <- hot_dog_model_final %>% fit(
#   train_generator,
#   steps_per_epoch = 198 / batch_size, # divide size of training dataset by batch size
#   epochs = 20,
#   validation_data = validation_generator,
#   validation_steps = 300 / batch_size, # divide size of training dataset by batch size
#   callback_reduce_lr_on_plateau(monitor = "val_loss", factor = 0.1) # if crossentropy loss plateaus, change LR
# )

# save fit
# saveRDS(hot_dog_final_aug3_RDS, paste0(path, "data/hot_dog_final_aug3_RDS.rds"))
hot_dog_final_aug3_RDS <- readRDS(paste0(path, "data/hot_dog_final_aug3_RDS.rds"))

# save model
# save_model_hdf5(hot_dog_model_final, paste0(path, "data/hot_dog_model_aug3.h5"))

# import model
hot_dog_model_final <- load_model_hdf5(paste0(path, "data/hot_dog_model_aug3.h5"))

# add performance to list
model_eval_list['third_try_augmentation'] <- as.data.frame(evaluate_generator(hot_dog_model_final, validation_generator, steps = 10))
```

```{r}
# turn performance list into dataframe
model_eval_hot_dog <- as.data.frame(model_eval_list)

# set row names
row.names(model_eval_hot_dog) <- c('loss','accuracy')

# set column names
colnames(model_eval_hot_dog) <- c('Final model (no aug.)', 'Model 1 aug.', 'Model 2 aug.', 'Model 3 aug.')

# create table
knitr::kable(model_eval_hot_dog, caption = "Performance comparison of augmented and non-augmented networks", digits = 3 ) %>% 
  kable_styling( position = "center", latex_options = 'hold_position', bootstrap_options = c("striped", "hover"))
```

### d. Try to rely on some pre-built neural networks to aid prediction. Can you achieve a better performance using transfer learning for this problem?

Even after applying augmentation techniques the performances of the models remained relatively low. So as a next step I used pre-trained networks to achieve a better performance. I also kept the best performing data augmentation settings. As for the pre-trained networks I used Inception V3 and Mobilenet. The first one was created by Google and has 21,802,784 parameters. The second one has 3,228,864 parameters. An important setting is to freeze the weights of these models so that they do not get recalculated during the training.

```{r, include=FALSE}
# set best data augmentation parameters
train_datagen = image_data_generator(
  rescale = 1/255,
  rotation_range = 40,
  width_shift_range = 0.2,
  height_shift_range = 0.2,
  shear_range = 0.2,
  zoom_range = 0.2, 
  horizontal_flip = TRUE,
  fill_mode = "nearest"
)

# reinitialize training generator
train_generator <- flow_images_from_directory(
  file.path(path, "data/hot-dog-not-hot-dog/train"), 
  train_datagen,              
  target_size = image_size,  
  batch_size = batch_size,
  class_mode = "binary"       
)
```

```{r, include=FALSE}
##### FIRST MODEL

# create the base pre-trained model
conv_base <- application_inception_v3(
  weights = "imagenet",
  include_top = FALSE,
  input_shape = c(image_size, 3)
)

# the pre-trained model has 21,802,784 parameters
summary(conv_base)

# freeze weights
freeze_weights(conv_base)

# add layers
hot_dog_model_final_inceptionV3 <- keras_model_sequential()%>%
  conv_base %>% 
  layer_global_average_pooling_2d() %>% 
  layer_dense(units = 16, activation = 'relu') %>% 
  layer_dense(units = 1, activation = 'sigmoid')

# compile model
hot_dog_model_final_inceptionV3 %>% compile(
  loss = "binary_crossentropy",
  optimizer = 'rmsprop',
  metrics = c("accuracy")
)

# fit model
# hot_dog_model_final_inceptionV3_RDS <- hot_dog_model_final_inceptionV3 %>% fit(
#   train_generator,
#   steps_per_epoch = 198 / batch_size, # divide size of training dataset by batch size
#   epochs = 10,
#   validation_data = validation_generator,
#   validation_steps = 300 / batch_size, # divide size of training dataset by batch size
#   callback_reduce_lr_on_plateau(monitor = "val_loss", factor = 0.1) # if crossentropy loss plateaus, change LR
# )

# save fit
# saveRDS(hot_dog_model_final_inceptionV3_RDS, paste0( path, "data/hot_dog_model_final_inceptionV3_RDS.rds"))
hot_dog_model_final_inceptionV3_RDS <- readRDS(paste0( path, "data/hot_dog_model_final_inceptionV3_RDS.rds"))

# save model
# save_model_hdf5(hot_dog_model_final_inceptionV3, paste0(path, "data/hot_dog_model_final_inceptionV3.h5"))

# import model
hot_dog_model_final_inceptionV3 <- load_model_hdf5(paste0(path, "data/hot_dog_model_final_inceptionV3.h5"))

# add model to list
model_eval_list['final_conv_inceptionV3_augmented'] <- 
  as.data.frame(evaluate_generator(hot_dog_model_final_inceptionV3, validation_generator, steps = 10))
```

```{r, include=FALSE}
##### SECOND TRY

# create the base pre-trained model
base_model <- application_mobilenet(
  weights = 'imagenet', 
  include_top = FALSE,
  input_shape = c(image_size, 3))

# the pre-trained model has 3,228,864 parameters
summary(base_model)

# freeze weights
freeze_weights(base_model)

# add layers
hot_dog_model_final_mobilenet <- keras_model_sequential()%>%
  base_model %>% 
  layer_global_average_pooling_2d() %>% 
  layer_dense(units = 16, activation = 'relu') %>% 
  layer_dense(units = 1, activation = 'sigmoid')

# compile model
hot_dog_model_final_mobilenet %>% compile(
  loss = "binary_crossentropy",
  optimizer = 'rmsprop',
  metrics = c("accuracy")
)

# fit model
# hot_dog_model_final_mobilenet_RDS <- hot_dog_model_final_mobilenet %>% fit(
#   train_generator,
#   steps_per_epoch = 198 / batch_size, # divide size of training dataset by batch size
#   epochs = 10,
#   validation_data = validation_generator,
#   validation_steps = 300 / batch_size, # divide size of training dataset by batch size
#   callback_reduce_lr_on_plateau(monitor = "val_loss", factor = 0.1) # if crossentropy loss plateaus, change LR
# )

# save fit
# saveRDS(hot_dog_model_final_mobilenet_RDS, paste0(path, "data/hot_dog_model_final_mobilenet_RDS.rds"))
hot_dog_model_final_mobilenet_RDS <- readRDS(paste0(path, "data/hot_dog_model_final_mobilenet_RDS.rds"))

# save model
# save_model_hdf5(hot_dog_model_final_mobilenet, paste0(path, "data/hot_dog_model_final_mobilenet.h5"))

# import model
hot_dog_model_final_mobilenet <- load_model_hdf5(paste0(path, "data/hot_dog_model_final_mobilenet.h5"))

# add model to list
model_eval_list['final_conv_mobilenet_augmented'] <- 
  as.data.frame(evaluate_generator(hot_dog_model_final_mobilenet, validation_generator, steps = 10))
```

The table below shows the validation set performances of all the image classifier models. We can conclude that data augmentation helped in achieving better performance but not for all augmented models. However, the most significant increase was achieved by using pre-trained networks.

```{r}
# final model comparison
# create data frame from model comparison list
model_eval_hot_dog <- as.data.frame(model_eval_list)

# rename rows
row.names(model_eval_hot_dog) <- c('loss','accuracy')

# set column names
colnames(model_eval_hot_dog) <- c('Final model (no aug.)', 'Model 1 aug.', 'Model 2 aug.', 'Model 3 aug.', 'Model inception v3', 'Model mobilenet')

# create table
knitr::kable(model_eval_hot_dog, caption = "Performance comparison of all image classifier models", digits = 3 ) %>% 
  kable_styling( position = "center", latex_options = 'hold_position', bootstrap_options = c("striped", "hover"))
```

